---
title: 'Portfolio building #2'
author: "Ahmed ALRashid"
date: "12/3/2020"
output: html_document
---


#Loading Packages
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggplot2)
library(dplyr)
library(caret)
library(recipes)
library(rsample)
library(class)
library(vip)
library(glmnet)
```


#Reading the data
```{r}
vg7auction <- read.csv("Xbox 7-day auctions.csv")
```

#Deleting missing values
```{r}
vg7auction <- na.omit(vg7auction)
```

#spliting the data
```{r}
set.seed(123)  
split  <- rsample::initial_split(vg7auction, prop = 0.7, 
                                 strata = "bid")
train_v  <- rsample::training(split)
test_v   <- rsample::testing(split)
```

#data prep
```{r}

# Create training  feature matrices
# we use model.matrix(...)[, -1] to discard the intercept
X <- model.matrix(bid ~ ., train_v)[, -1]

# transform y with log transformation
Y <- log(train_v$bid)
```



1. Depending on the type of response variable, apply a linear or logistic regression model.
First, apply the model to your data without pre-applying feature engineering processes.
```{r}

model1 <- lm(bid ~ price, data = train_v)
summary(model1)
# RMSE
sigma(model1)
# MSE
sigma(model1)^2
confint(model1, level = 0.95)

```


Create and a apply a blueprint of feature engineering processes that you think will help your model improve.

```{r}

blueprint <- recipe(bid ~ ., data = train_v) %>%
  step_nzv(all_nominal())  %>%
  step_center(all_numeric(), -all_outcomes()) %>%
  step_scale(all_numeric(), -all_outcomes()) %>%
  step_pca(all_numeric(), -all_outcomes())
  
blueprint

```

```{r}
prepare <- prep(blueprint, training = train_v)
prepare
```


Now reapply the model to your data that has been feature engineered.
Did your model performance improve?
No
```{r}

model1 <- lm(bid ~ price, data = train_v)
summary(model1)
# RMSE
sigma(model1)
# MSE
sigma(model1)^2
confint(model1, level = 0.95)

```


2. Apply a principal component regression model.

```{r}

set.seed(123)
cv_model_pcr <- train(
  price ~ ., 
  data = train_v, 
  method = "pcr",
  trControl = trainControl(method = "cv", number = 10),
  preProcess = c("zv", "center", "scale"),
  tuneLength = 100
  )

# model with lowest RMSE
cv_model_pcr$bestTune

```

Perform a grid search over several components.
Identify and explain the performance of the optimal model.



3.Apply a partial least squares regression model.

```{r}

# perform 10-fold cross validation on a PLS model tuning the 
# number of principal components to use as predictors from 1-30
set.seed(123)
cv_model_pls <- train(
  bid ~ ., 
  data = train_v, 
  method = "pls",
  trControl = trainControl(method = "cv", number = 10),
  preProcess = c("zv", "center", "scale"),
  tuneLength = 30
)

# model with lowest RMSE
cv_model_pls$bestTune

```



Perform a grid search over several components.
```{r}

# for reproducibility
set.seed(123)

# grid search across 
cv_glmnet <- train(
  x = X,
  y = Y,
  method = "glmnet",
  preProc = c("zv", "center", "scale"),
  trControl = trainControl(method = "cv", number = 10),
  tuneLength = 10
)

# model with lowest RMSE
cv_glmnet$bestTune

```

Identify and explain the performance of the optimal model.



4. Apply a regularized regression model.

Ridge:
```{r}
ridge <- glmnet(
  x = X,
  y = Y,
  alpha = 0
)

plot(ridge, xvar = "lambda")
```

Lasso:
```{r}
lasso <- glmnet(
  x = X,
  y = Y,
  alpha = 1
)

plot(lasso, xvar = "lambda")
```

Ridge CV model:
```{r}
ridge <- cv.glmnet(
  x = X,
  y = Y,
  alpha = 0
)

plot(ridge)
```

Lasso CV model:
```{r}
lasso <- cv.glmnet(
  x = X,
  y = Y,
  alpha = 1
)

plot(lasso)
```

Ridge model results: 
```{r}
# Ridge model - minimum MSE
min(ridge$cvm)

# Ridge model - lambda for this min MSE
ridge$lambda.min 

# Ridge model w/1-SE rule
ridge$cvm[ridge$lambda == ridge$lambda.1se]

# Ridge model w/1-SE rule -- No. of coef | 1-SE MSE
ridge$nzero[ridge$lambda == ridge$lambda.1se]
```

Lasso model results: 
```{r}
# Lasso model - minimum MSE
min(lasso$cvm)       

# Lasso model - lambda for this min MSE
lasso$lambda.min 

# Lasso model - w/1-SE rule
lasso$cvm[lasso$lambda == lasso$lambda.1se]

# Lasso model w/1-SE rule -- No. of coef | 1-SE MSE
lasso$nzero[lasso$lambda == lasso$lambda.1se]
```




Perform a grid search across alpha parameter values ranging between 0–1.

```{r cv-glmnet}
# tuning grid
hyper_grid <- expand.grid(
  alpha = seq(0, 1, by = .25),
  lambda = c(0.1, 10, 100, 1000, 10000)
)

# perform resampling
set.seed(123)
cv_glmnet <- train(
  x = X,
  y = Y,
  method = "glmnet",
  preProc = c("zv", "center", "scale"),
  trControl = trainControl(method = "cv", number = 10),
  tuneLength = 10
)

# best model
cv_glmnet$results %>%
  filter(
    alpha == cv_glmnet$bestTune$alpha,
    lambda == cv_glmnet$bestTune$lambda
  )
```

```{r cv-glmnet-plot, fig.height=5}
# plot results
plot(cv_glmnet)
```


What is the optimal alpha and lambda values?
What is the MSE and RMSE for this optimal model?
How does it compare to your previous models?

```{r}
# predict sales price on training data
pred <- predict(cv_glmnet, X)

# compute RMSE of transformed predicted
RMSE(exp(pred), exp(Y))
```


5. Pick the best performing model from above.
Identify the most influential features for this model.

# Feature importance
```{r}
vip(cv_glmnet, num_features = 20, geom = "point")
```
Plot the top 10 most influential features.
Do these features have positive or negative impacts on your response variable?
